White Dwarf Mergers on Adaptive Meshes\\ II. Nuclear Reactions

### 2017-11-14 papers/wdmerger_II/paper.tex
Move the timestep limiter discussion into the methodology section

@@

%==========================================================================
% Temporal Resolution
%==========================================================================
\section{Temporal Resolution}
\label{sec:temporalresolution}

For the standard self-heating burns that we use, the timestep limiting scheme
described in \autoref{sec:hydrocoupling} has a significant effect on the evolution.
The parameters $f_{be}$ (from \autoref{eq:timestep_e}) and $f_{bX}$ (from
\autoref{eq:timestep_X}) control the size of the timestep as a function of the
burning rate. As the timestep gets smaller, the error due to the Strang splitting
scheme coupling the reactions and the hydro also gets smaller, and we expect
that the results become more accurate.
\begin{figure}[ht]
  \centering
  \includegraphics[scale=0.425]{{{plots/dtnuc_max_Ni56_m_P_0.64_m_S_0.64}}}
  \caption{Nickel production as a function of the timestep factors $f_{be}$
           (solid blue, limiting the timestep on changes in internal energy) and
           $f_{bX}$ (dashed green, limiting the timestep on changes in species).
           For a given $f_{be}$, $f_{bX}$ with the same value typically limits
           the timestep by at least an order of magnitude more.
           \label{fig:timestep_factor}}
\end{figure}
\autoref{fig:timestep_factor} demonstrates the effect of both parameters on the nickel
production (these limiters are somewhat redundant in the sense that a given $f_{bX}$
roughly corresponds to another (higher) $f_{be}$, as is evident by the fact that on
the graph the curves look qualitatively similar in shape but with a horizontal offset.
As either $f$ becomes small, the restriction on the timestep becomes quite severe:
for $f_{be} = 0.001$, the minimum timestep in the simulation is smaller than
$5 \times 10^{-9}\ \text{s}$ and the full simulation requires over 100,000 timesteps.
Such a simulation would be prohibitively expensive for a high resolution 3D calculation.
It is clear, though, that the timestep effect has converged when $f_{be} \sim 0.1$. We
agree that the use of values in this range by \cite{raskin:2010}, \cite{hawley:2012},
and \cite{kushnir:2013} is appropriate.

>>>>

\begin{figure}[ht]
  \centering
  \includegraphics[scale=0.425]{{{plots/dtnuc_max_Ni56_m_P_0.64_m_S_0.64}}}
  \caption{Nickel production as a function of the timestep factors $f_{be}$
           (solid blue, limiting the timestep on changes in internal energy) and
           $f_{bX}$ (dashed green, limiting the timestep on changes in species).
           For a given $f_{be}$, $f_{bX}$ with the same value typically limits
           the timestep by at least an order of magnitude more.
           \label{fig:timestep_factor}}
\end{figure}
\autoref{fig:timestep_factor} demonstrates the effect of both parameters on the nickel
production for the baseline collision case. These limiters are somewhat redundant in
the sense that a given $f_{bX}$ roughly corresponds to another (higher) $f_{be}$, as
is evident by the fact that on the plot the curves look qualitatively similar in
shape but with a horizontal offset. As either $f$ becomes small, the restriction on
the timestep becomes more severe: for $f_{bX} = 0.1$, the full simulation requires
nearly 30,000 timesteps, compared to just over 1,000 for the baseline case.
Such a simulation would be quite expensive for a high resolution 3D calculation.
It is clear, though, that even though the simulation is not converged when the
timestep is not limited due to burning, the effect is small in absolute terms.
So we agree that the use of values in this range by \cite{raskin:2010},
\cite{hawley:2012}, and \cite{kushnir:2013} is appropriate. However, for reasons of
computational expense we do not use the timestep limiter in the high resolution
simulations that follow. The error due to this choice should be small, because
the effect is not that large to begin with, and because the higher resolution
decreases the timestep anyway due to the hydrodynamic stability constraint.

@@
